GVN-Hoist: Hoisting Computations from Branches

### 2016-08-23 hoist.tex
reword some more

\subsection{Checking for legality}
\label{subsec:legality}
Since the equality of candidates is purely based on the value they compute, we
need to establish if hoisting them to a common dominator would be feasible. One
a common dominator is found then we check if all the use-operands of the set of
instructions are available or not. In some cases when the operands are not
available, it is possible to reinstantiate (remateralize) the use-operands, thus
passing the legality check. Subsequently, it is checked that the side-effects of
the computations does not intersect with any side-effects between the
instructions to be hoisted and their hoisting point. For memory operations like
loads, stores, calls etc., it is also required to check that all the paths from
the hoisting point to the end of the function should execute the exact
instruction, in order to guarantee correctness. Moreover, hoisting of memory
operations is tricky on paths which have indirect branch targets e.g., landing
pad, case statements, goto labels etc., because it becomes difficult to prove
that all the paths from hoisting point to the end of the function would execute
the instruction. In our current implementation we discard hoisting through such
paths.

In the optimistic approach (Section-\ref{subsec:optimistic}), it is possible
that a common hoisting point of all the instructions is either too far away, or
not legally possible. In these cases, it is still possible to `partially' hoist
a subset of instructions by splitting the set of candidates and finding a closer
hoisting point for each subset. For more details see
Section-\ref{subsec:partition}.

\subsubsection{Hoisting scalars}
Scalars are the easiest to hoist because we do not have to analyze them for
aliasing memory references. As long as all the operands are available (or can be
made available by rematerialization), the scalar can be hoisted.

\subsubsection{Hoisting loads}
The availability of operand to the load (an address) is checked at the hoisting
point. If that is not available we try to rematerialize the addresss if
possible.  Along the path, from current position of the load instruction through
the hoisting point, we find if there are aliasing writes to memory. In those
cases we discard the candidate.

\subsubsection{Hoisting stores}
We need to check the dependency requirements similar to the hoisting of
loads. We need to check that the operands of the store instruction are available
at the hoisting point, that there are no aliasing loads or store along the path
from the current position through the hoisting point.

\subsubsection{Hoisting calls}
For hoisting the call instructions, we check that the loads/stores in the call
would not hazard with any loads/stores (in the caller) along the path from the
callsite through the hoisting point, just like we did for loads and stores.

>>>>

\subsection{Checking for legality}
\label{subsec:legality}
Since the equality of candidates is purely based on the value they compute, we
need to establish if hoisting them to a common dominator would be feasible. Once
a common dominator is found, we check whether all the use-operands of the set of
instructions are available at that position. In some cases when the operands are not
available, it is possible to reinstantiate (remateralize) the use-operands, thus
passing the legality check.

Subsequently, it is checked that the side-effects of the computations does not
intersect with any side-effects between the instructions to be hoisted and their
hoisting point. For memory operations like loads, stores, calls etc., it is also
required to check that all the paths from the hoisting point to the end of the
function should execute the exact instruction, in order to guarantee
correctness.

Moreover, hoisting of memory operations is tricky on paths which have indirect
branch targets e.g., landing pad, case statements, goto labels etc., because it
becomes difficult to prove that all the paths from hoisting point to the end of
the function would execute the instruction. In our current implementation we
discard hoisting through such paths.

In the optimistic approach, described in Section-\ref{subsec:optimistic}, it is
possible that a common hoisting point of all the instructions is either too far
away, or not legally possible. In these cases, it is still possible to
`partially' hoist a subset of instructions by splitting the set of candidates
and finding a closer hoisting point for each subset. For more details see
Section-\ref{subsec:partition}.

\subsubsection{Hoisting scalars}
Scalars are the easiest to hoist because we do not have to analyze them for
aliasing memory references. As long as all the operands are available (or can be
made available by rematerialization), the scalar computation can be hoisted.

\subsubsection{Hoisting loads}
The availability of operand to the load (an address) is checked at the hoisting
point. If that is not available we try to rematerialize the addresss if
possible.  Along the path, from current position of the load instruction
backwards on the control flow to the hoisting point, we check whether there are
writes to memory that may alias with the load, in which case we discard the
candidate.

\subsubsection{Hoisting stores}
For stores, we check the dependency requirements similar to the hoisting of
loads. We check that the operands of the store instruction are available at the
hoisting point, that there are no aliasing loads or store along the path from
the current position to the hoisting point.

\subsubsection{Hoisting calls}
Call instructions can be divided into three categories: those calls equivalent
to purely scalar computations, calls reading from memory, and most of the time,
without further information, calls have to be classified as writing to memory,
that is the most restrictive form.  Each category of call instructions will be
handled as described for scalar, load, and store instructions.
